---
title: "Working with Geodata in R"
subtitle: "E1493 - Data Journalism"
author: "Simon Munzert"
institution: "Hertie School"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: false
    theme: cosmo
    code-fold: false
    code-tools: true
    df-print: paged
    fig-width: 8
    fig-height: 5
    highlight-style: github
    embed-resources: true
    link-external-newwindow: true
execute:
  eval: false
  warning: false
  message: false
---

```{r}
#| eval: true
#| include: false
#| label: load-packages

pkgs <- c("tidyverse", "sf", "terra", "rnaturalearth",
          "rnaturalearthdata", "nominatimlite", "tidygeocoder")
lapply(pkgs, library, character.only = TRUE)
```

::: {.callout-note}
## Before you start

Install the packages used in this workshop:

```r
install.packages(c(
  "tidyverse", "sf", "terra", "stars",
  "rnaturalearth", "rnaturalearthdata", "rnaturalearthhires",
  "nominatimlite", "tidygeocoder", "osmdata"
))
```

This script accompanies the **Session 6 slides** on Mapping and Geodata. It's designed as a hands-on complement — the slides cover concepts like spatial data types, projections, and distance; this script lets you work with real spatial data in R. We use the **German breweries** dataset throughout as a running example.

The script builds on concepts from Pebesma & Bivand (2023), [*Spatial Data Science: With Applications in R*](https://r-spatial.org/book/), freely available online. If you want to go deeper, that book is the definitive reference for `sf` and `stars`.
:::

---

# Part 1: Importing Spatial Data {#importing}

The two main types of spatial data — **vector** (points, lines, polygons) and **raster** (gridded pixels) — each have their own packages and workflows in R. The `sf` package handles vector data; `terra` (and `stars`) handle raster data. Both integrate with the tidyverse.


## Vector data with `sf`

The workhorse function is `st_read()`, which can read shapefiles, GeoJSON, GeoPackage, KML, and dozens of other formats via GDAL. Let's start by loading a polygon map of Germany from the `rnaturalearth` package:

```{r}
#| eval: true
#| label: import-germany

library(sf)
library(rnaturalearth)

# Country boundary as an sf object
germany <- ne_countries(
  scale = "medium",
  country = "Germany",
  returnclass = "sf"
)

# What does an sf object look like?
class(germany)
glimpse(germany)
```

An `sf` object is just a data frame with a special `geometry` column. This means all your familiar `dplyr` verbs work on it — `filter()`, `mutate()`, `select()`, `left_join()`, you name it. The geometry just comes along for the ride.

```{r}
#| eval: true
#| label: sf-structure

# The geometry column
st_geometry(germany)

# CRS (coordinate reference system)
st_crs(germany)
```

The CRS tells you how coordinates map onto the Earth's surface. Here we have WGS 84 (EPSG:4326) — the standard longitude/latitude system used by GPS.


## Reading files from disk

In real projects, you'll often read shapefiles or GeoJSON files. The syntax is the same regardless of format — `st_read()` detects the driver automatically:

```{r}
#| label: read-shapefile

# Reading a shapefile (hypothetical example)
# districts <- st_read("data/german_districts.shp")

# Reading GeoJSON
# districts <- st_read("data/german_districts.geojson")

# Reading GeoPackage (modern, recommended format)
# districts <- st_read("data/german_districts.gpkg")
```

::: {.callout-tip}
## Shapefile vs. GeoPackage

Shapefiles are the legacy standard — they consist of 4-6 files (`.shp`, `.dbf`, `.shx`, `.prj`, ...) that must travel together. GeoPackage (`.gpkg`) is the modern replacement: a single file, no column name truncation, no file size limits. Use GeoPackage when you can.
:::


## Getting German state boundaries

For our brewery analysis, we'll need state-level polygons. Let's get them:

```{r}
#| eval: true
#| label: german-states

# German federal states (Bundesländer)
germany_states <- ne_states(country = "Germany", returnclass = "sf")

ggplot(germany_states) +
  geom_sf(fill = "gray95", color = "gray40") +
  theme_minimal(base_size = 14) +
  labs(title = "German federal states (Bundesländer)")
```

::: {.callout-note}
## Where to find spatial data

Good sources for administrative boundaries and other geodata include:

- [Natural Earth](https://www.naturalearthdata.com/) — global boundaries at multiple scales (via `rnaturalearth`)
- [GADM](https://gadm.org/) — detailed subnational boundaries
- [Geoportal.de](https://www.geoportal.de/) — official German geodata
- [OpenStreetMap](https://www.openstreetmap.org/) — crowd-sourced, very detailed (via `osmdata`)
- [Eurostat GISCO](https://ec.europa.eu/eurostat/web/gisco) — European administrative boundaries
:::


## Raster data with `terra`

Raster data represents continuous surfaces — elevation, temperature, satellite imagery — as regular grids of pixels. The `terra` package (successor to `raster`) is the standard tool:

```{r}
#| eval: true
#| label: raster-intro

library(terra)

# Create a simple example raster covering Germany
# (In practice, you'd read a real file with rast())
r <- rast(
  xmin = 5.5, xmax = 15.5,
  ymin = 47, ymax = 55.5,
  resolution = 0.1,
  crs = "EPSG:4326"
)

# Fill with random "elevation" values for demonstration
values(r) <- runif(ncell(r), min = 0, max = 2000)
r
```

```{r}
#| eval: true
#| label: raster-plot

# Quick plot
plot(r, main = "Simulated raster over Germany")
```

In real projects, you'd read raster files like GeoTIFF or NetCDF:

```{r}
#| label: raster-read

# Reading real raster data
# elevation <- rast("data/germany_dem.tif")
# temperature <- rast("data/temperature_2024.nc")
```

::: {.callout-tip}
## `terra` vs. `stars`

Both packages handle raster data. `terra` is faster and more memory-efficient for classic raster operations. `stars` integrates better with `sf` and handles data cubes (e.g., space × time). For most data journalism work, `terra` is the pragmatic choice.
:::

---

# Part 2: Projections {#projections}

A coordinate reference system (CRS) defines how coordinates map onto the Earth. Getting this right matters — wrong projections lead to distorted maps and incorrect distance/area calculations.


## The key distinction: geographic vs. projected

**Geographic CRS** (e.g., WGS 84 / EPSG:4326): Coordinates are in degrees of longitude and latitude on an ellipsoidal model of the Earth. This is what GPS uses.

**Projected CRS** (e.g., UTM Zone 32N / EPSG:32632): Coordinates are in meters (or feet) on a flat 2D plane. Every projection introduces some distortion — the question is which property you want to preserve: area, shape, distance, or direction.


## Checking and setting CRS

```{r}
#| eval: true
#| label: check-crs

# Check the CRS of our Germany polygon
st_crs(germany)

# The EPSG code is the shorthand
st_crs(germany)$epsg
```


## Reprojecting data

Use `st_transform()` to reproject vector data. For Germany, a good projected CRS is **ETRS89 / UTM Zone 32N** (EPSG:25832) — this is the official German reference system:

```{r}
#| eval: true
#| label: reproject

# Reproject to ETRS89 / UTM zone 32N
germany_proj <- st_transform(germany, crs = 25832)

# Now coordinates are in meters, not degrees
st_bbox(germany)       # in degrees
st_bbox(germany_proj)  # in meters
```

```{r}
#| eval: true
#| label: projection-comparison
#| fig-height: 4

# Visual comparison
p1 <- ggplot(germany) +
  geom_sf(fill = "gray90") +
  labs(title = "WGS 84 (geographic)") +
  theme_minimal()

p2 <- ggplot(germany_proj) +
  geom_sf(fill = "gray90") +
  labs(title = "ETRS89 / UTM 32N (projected)") +
  theme_minimal()

library(patchwork)
p1 + p2
```

The visual difference for Germany is subtle, but the implications for calculations are real: area and distance computations on projected data use planar geometry (fast, exact), while geographic coordinates require spherical/ellipsoidal calculations.


## When does the projection matter?

For **display** (making maps), the projection choice is mostly aesthetic. For **analysis** (calculating distances, areas, buffers), it's critical. A common workflow:

1. Store and exchange data in WGS 84 (EPSG:4326) — the universal standard.
2. Reproject to an appropriate local CRS before doing any spatial calculations.
3. Reproject back to WGS 84 for web mapping or data exchange if needed.

```{r}
#| eval: true
#| label: area-comparison

# Area of Germany in different CRS
st_area(germany) |> units::set_units("km^2")        # spherical calculation
st_area(germany_proj) |> units::set_units("km^2")    # planar calculation
```

::: {.callout-important}
## A common trap

If you compute distances or areas on unprojected (lon/lat) data, `sf` will use spherical geometry by default (via the `s2` library), which is correct but slower. If you accidentally treat lon/lat as planar coordinates (e.g., by disabling `s2`), you'll get wrong results — especially at higher latitudes where one degree of longitude is much less than 111 km.
:::

---

# Part 3: Transformations and Spatial Overlay {#transformations}

Spatial transformations let you combine datasets: which breweries are in Bavaria? How many points fall within each state? What's the nearest city to each brewery? This is where spatial data becomes powerful for journalism.


## Loading our brewery data

Let's load the geocoded brewery coordinates from the scraping exercise:

```{r}
#| eval: true
#| label: load-breweries

# Load the pre-geocoded brewery data
load("coords_breweries.RData")
head(coords_df)

# Clean: filter to plausible German coordinates
breweries <- coords_df |>
  filter(
    !is.na(lon), !is.na(lat),
    lon >= 5.5, lon <= 15.5,
    lat >= 47, lat <= 55.5
  )

nrow(breweries)
```


## Converting a data frame to an `sf` object

The brewery data is just a regular data frame with lon/lat columns. To make it spatial, we use `st_as_sf()`:

```{r}
#| eval: true
#| label: breweries-sf

breweries_sf <- st_as_sf(
  breweries,
  coords = c("lon", "lat"),
  crs = 4326  # WGS 84
)

# It's now a spatial object
class(breweries_sf)
head(breweries_sf)
```

```{r}
#| eval: true
#| label: breweries-quick-map

# Quick check: plot breweries on top of Germany
ggplot() +
  geom_sf(data = germany, fill = "gray95") +
  geom_sf(data = breweries_sf, color = "red", size = 0.4, alpha = 0.5) +
  theme_minimal(base_size = 14) +
  labs(title = "Breweries in Germany")
```


## Spatial join: which state is each brewery in?

The classic overlay operation: assign each point to the polygon it falls within. This is `st_join()` — the spatial equivalent of `left_join()`:

```{r}
#| eval: true
#| label: spatial-join

# Join breweries to states: each brewery gets the attributes of the state it's in
breweries_with_state <- st_join(breweries_sf, germany_states["name"])

head(breweries_with_state)
```

```{r}
#| eval: true
#| label: breweries-per-state

# Count breweries per state
breweries_per_state <- breweries_with_state |>
  st_drop_geometry() |>
  count(name, sort = TRUE) |>
  rename(state = name, n_breweries = n)

breweries_per_state
```

No surprise: Bavaria dominates. Let's put this on a map.

```{r}
#| eval: true
#| label: choropleth-breweries
#| fig-height: 6

# Merge counts back to the state polygons
states_with_breweries <- germany_states |>
  left_join(breweries_per_state, by = c("name" = "state"))

ggplot(states_with_breweries) +
  geom_sf(aes(fill = n_breweries), color = "white") +
  scale_fill_viridis_c(
    option = "inferno",
    direction = -1,
    na.value = "gray90",
    name = "Breweries"
  ) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Number of breweries by German state",
    caption = "Source: biermap24.de"
  )
```


## Buffering, intersection, and other geometric operations

`sf` provides a full set of geometric operations. Here are some commonly useful ones:

```{r}
#| eval: true
#| label: geometric-ops

# Reproject for metric operations
breweries_proj <- st_transform(breweries_sf, 25832)
germany_states_proj <- st_transform(germany_states, 25832)

# Buffer: 50 km circle around Berlin
berlin_point <- st_sfc(st_point(c(13.405, 52.52)), crs = 4326) |>
  st_transform(25832)
berlin_buffer <- st_buffer(berlin_point, dist = 50000)  # 50 km in meters

# Which breweries fall within 50 km of Berlin?
breweries_near_berlin <- st_intersection(breweries_proj, berlin_buffer)
cat("Breweries within 50 km of Berlin:", nrow(breweries_near_berlin), "\n")
```

```{r}
#| eval: true
#| label: buffer-map

# Visualize the buffer
ggplot() +
  geom_sf(data = germany_states_proj, fill = "gray95", color = "gray60") +
  geom_sf(data = berlin_buffer, fill = "steelblue", alpha = 0.2, color = "steelblue") +
  geom_sf(data = breweries_proj, color = "red", size = 0.3, alpha = 0.3) +
  geom_sf(data = breweries_near_berlin, color = "darkred", size = 1.5) +
  coord_sf(
    xlim = st_bbox(berlin_buffer)[c(1, 3)] + c(-100000, 100000),
    ylim = st_bbox(berlin_buffer)[c(2, 4)] + c(-100000, 100000)
  ) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Breweries within 50 km of Berlin",
    subtitle = "Blue circle = 50 km buffer around city center"
  )
```


## Key transformation functions at a glance

Here are the `sf` functions you'll use most often for spatial analysis:

```{r}
#| eval: false
#| label: transformation-cheatsheet

# Predicates (return TRUE/FALSE or index lists)
st_intersects(x, y)     # do geometries intersect?
st_within(x, y)         # is x fully within y?
st_contains(x, y)       # does x fully contain y?
st_is_within_distance(x, y, dist)  # within distance threshold?

# Operations (return new geometries)
st_intersection(x, y)   # clip x to y (overlay)
st_difference(x, y)     # x minus y
st_union(x)             # merge all geometries into one
st_buffer(x, dist)      # expand/contract by distance
st_centroid(x)          # center point of polygons

# Measurements
st_distance(x, y)       # pairwise distances
st_area(x)              # polygon areas
st_length(x)            # line lengths

# Joins
st_join(x, y)           # spatial left join (like dplyr::left_join, but by geometry)
st_nearest_feature(x, y) # index of nearest feature in y for each x
```

::: {.callout-note}
## For data journalists

The spatial join (`st_join`) is your single most useful function. It lets you answer questions like: "Which district is this event in?", "How many incidents occurred in each neighborhood?", "Which schools are in the flood zone?" — all without manual matching or lookup tables.
:::

---

# Part 4: Geocoding {#geocoding}

Geocoding turns addresses or place names into coordinates. It's essential when your source data contains location names but no geometry — think lists of schools, hospitals, crime incidents, or... breweries.


## Option 1: `tidygeocoder` (recommended for most use cases)

`tidygeocoder` is a unified interface to multiple geocoding services. It works directly on data frames and integrates cleanly with the tidyverse. The **Nominatim** (OpenStreetMap) backend is free and requires no API key:

```{r}
#| eval: true
#| label: tidygeocoder-demo

library(tidygeocoder)

# Geocode a few German cities
cities <- tibble(
  name = c("München", "Hamburg", "Köln", "Frankfurt am Main", "Stuttgart")
)

cities_geocoded <- cities |>
  geocode(
    address = name,
    method = "osm"  # Nominatim / OpenStreetMap — free, no key needed
  )

cities_geocoded
```

The method parameter lets you switch between services. Here's what's available for free:

```{r}
#| label: tidygeocoder-services

# Free services (no API key required):
# method = "osm"        — Nominatim/OpenStreetMap. Best for international use.
#                          Rate limit: 1 request/second.
# method = "arcgis"     — ArcGIS. Good international coverage, no key for basic use.
# method = "census"     — US Census. US addresses only.

# Paid services (API key required, but often have free tiers):
# method = "google"     — Google Maps. Excellent quality, $5/1000 requests.
# method = "mapbox"     — Mapbox. 100k free requests/month.
# method = "geocodio"   — Geocodio. 2,500 free/day. US & Canada only.
# method = "here"       — HERE. 250k free requests/month.
```

::: {.callout-tip}
## My recommendation

For most data journalism projects, **Nominatim** (via `method = "osm"`) is the right starting point: it's free, requires no API key, and has good global coverage. Its main limitation is the rate limit (1 request/second), which makes it slow for large batches. For large-scale geocoding (10,000+ addresses), consider **ArcGIS** (also free, faster) or a paid service like **Mapbox** or **HERE**.
:::


## Option 2: `nominatimlite` (lightweight, sf-native)

If you want geocoded results directly as `sf` objects (instead of lat/lon columns), `nominatimlite` is a nice lightweight option:

```{r}
#| eval: true
#| label: nominatimlite-demo

library(nominatimlite)

# Geocode and get sf geometry directly
munich_sf <- geo_lite_sf("München, Germany", limit = 1)
munich_sf

# Geocode multiple places
brewery_cities <- c("Bamberg, Germany", "Kulmbach, Germany", "Bayreuth, Germany")
brewery_cities_sf <- geo_lite_sf(brewery_cities, limit = 1)
brewery_cities_sf
```


## Geocoding the brewery dataset

In the `scraping-breweries.R` script, we geocoded brewery cities using the Nominatim API. Here's what that workflow looks like with `tidygeocoder` — a cleaner, more modern approach:

```{r}
#| label: geocode-breweries-modern

# If you needed to re-geocode the brewery cities from scratch:
# (Don't run this — it takes a while due to rate limits)

# unique_cities <- unique(cities_from_scrape)
#
# geocoded <- tibble(city = unique_cities) |>
#   geocode(
#     address = city,
#     method = "osm",
#     custom_query = list(countrycodes = "de")  # restrict to Germany
#   )
```

::: {.callout-important}
## Geocoding gotchas

1. **Ambiguity.** "Frankfurt" could be Frankfurt am Main or Frankfurt an der Oder. Always add country context (e.g., `countrycodes = "de"`).
2. **Rate limits.** Nominatim allows 1 request/second. For 500+ cities, that's 8+ minutes. `tidygeocoder` handles rate limiting automatically.
3. **Quality varies.** Geocoding works well for cities and street addresses, but poorly for vague locations ("near the old bridge"). Always sanity-check results on a map.
4. **Cache your results.** Geocoding is slow and API-dependent. Save results to a file (`save()` or `write_csv()`) so you don't re-geocode every time.
:::


## Reverse geocoding

Sometimes you have coordinates and need addresses — that's reverse geocoding:

```{r}
#| eval: true
#| label: reverse-geocode

# What's at these coordinates?
coords <- tibble(
  lat = c(52.5200, 48.1351, 50.1109),
  lon = c(13.4050, 11.5820, 8.6821)
)

coords |>
  reverse_geocode(
    lat = lat, long = lon,
    method = "osm",
    address = "address_found"
  )
```

Reverse geocoding is useful when you have GPS traces, satellite data, or geotagged photos and need to identify the location in human terms.

---

# Part 5: Plotting Maps {#plotting}

Map design matters in data journalism — a well-crafted map can reveal patterns that tables and bar charts can't, but a bad map can mislead. This section covers practical map-making in R with an eye toward journalistic standards.


## The basic recipe: `ggplot2` + `geom_sf()`

Since `sf` objects are data frames, they plug directly into `ggplot2` via `geom_sf()`. This is the single most important thing to know about mapping in R:

```{r}
#| eval: true
#| label: basic-map
#| fig-height: 6

ggplot() +
  geom_sf(data = germany_states, fill = "gray95", color = "gray40") +
  geom_sf(data = breweries_sf, color = "#C2001A", size = 0.3, alpha = 0.4) +
  theme_void(base_size = 14) +
  labs(
    title = "Breweries in Germany",
    caption = "Source: biermap24.de"
  )
```

Already not bad. But let's make it better.


## Choropleth maps: encoding data in color

The most common journalistic map type. The key decisions are: what variable to show, which color scale, and how to bin the data.

```{r}
#| eval: true
#| label: choropleth-polished
#| fig-height: 6

# Calculate brewery density (per 100k population would be better,
# but we'll use per state for now)
states_map <- states_with_breweries |>
  mutate(n_breweries = replace_na(n_breweries, 0))

ggplot(states_map) +
  geom_sf(aes(fill = n_breweries), color = "white", linewidth = 0.3) +
  scale_fill_distiller(
    palette = "YlOrRd",
    direction = 1,
    name = "Number of\nbreweries",
    breaks = c(0, 50, 100, 200, 400, 600)
  ) +
  theme_void(base_size = 14) +
  theme(
    legend.position = c(0.15, 0.5),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = "gray40")
  ) +
  labs(
    title = "Germany's brewery landscape",
    subtitle = "Bavaria dominates with more than half of all breweries",
    caption = "Source: biermap24.de"
  )
```

::: {.callout-warning}
## The choropleth trap

Choropleth maps encode data by *area*, which means large but sparsely populated regions dominate visually. A map of "total breweries per state" will be dominated by Bavaria partly because Bavaria is *big*. Always consider normalizing: breweries per capita, per km², or using a different map type entirely.
:::


## Dot density maps: showing individual locations

When you have point data, showing the actual points is often more honest than aggregating into choropleth bins:

```{r}
#| eval: true
#| label: dot-density
#| fig-height: 6

ggplot() +
  geom_sf(data = germany_states, fill = "#f5f5f0", color = "gray70",
          linewidth = 0.2) +
  geom_sf(data = breweries_sf, color = "#C2001A", size = 0.15,
          alpha = 0.35) +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  labs(
    title = "Every brewery in Germany",
    subtitle = "Each dot = one brewery. The Franconian cluster is clearly visible.",
    caption = "Source: biermap24.de"
  )
```

This is the approach used in the slide image. The clustering in Franconia (northern Bavaria) is immediately obvious — something a choropleth at the state level would hide.


## Combining layers: points + polygons + labels

Journalistic maps often combine multiple layers. Let's highlight Bavaria and add context:

```{r}
#| eval: true
#| label: layered-map
#| fig-height: 6

# Isolate Bavaria
bavaria <- germany_states |> filter(name == "Bayern")

# Major cities for context
major_cities <- tibble(
  name = c("München", "Berlin", "Hamburg", "Bamberg", "Nürnberg"),
  lon = c(11.58, 13.41, 9.99, 10.89, 11.08),
  lat = c(48.14, 52.52, 53.55, 49.89, 49.45)
) |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

ggplot() +
  # Base: all states in light gray
  geom_sf(data = germany_states, fill = "#f0f0f0", color = "gray70",
          linewidth = 0.2) +
  # Highlight Bavaria
  geom_sf(data = bavaria, fill = "#FFF3CD", color = "#B8860B",
          linewidth = 0.4) +
  # All breweries
  geom_sf(data = breweries_sf, color = "#C2001A", size = 0.15,
          alpha = 0.35) +
  # Major cities as labeled points
  geom_sf(data = major_cities, color = "black", size = 2, shape = 21,
          fill = "white") +
  ggrepel::geom_text_repel(
    data = major_cities |> mutate(
      lon = st_coordinates(major_cities)[, 1],
      lat = st_coordinates(major_cities)[, 2]
    ),
    aes(x = lon, y = lat, label = name),
    size = 3.2, fontface = "bold",
    nudge_y = 0.3, segment.size = 0.3
  ) +
  theme_void(base_size = 14) +
  theme(plot.title = element_text(face = "bold", size = 16)) +
  labs(
    title = "Germany's breweries cluster in Bavaria and Franconia",
    caption = "Source: biermap24.de"
  )
```


## Best practices for journalistic maps

Here are concrete guidelines drawn from data journalism practice:

**1. Choose the right map type for the story.**
Point maps for discrete events (crimes, businesses, incidents). Choropleth maps for rates or aggregated data (unemployment rate, vote share). Avoid choropleths for raw counts — normalize first.

**2. Use sequential color scales for quantitative data.**
Single-hue scales (e.g., light to dark blue) work for "more vs. less." Diverging scales (e.g., blue–white–red) work when there's a meaningful midpoint (e.g., above/below average). Never use rainbow scales — they have no perceptual ordering.

**3. Keep it minimal.**
Strip away grid lines, axis labels, and unnecessary borders. Use `theme_void()` as your starting point. Add labels only where they serve the reader.

**4. Mind the visual area bias.**
Large regions attract attention regardless of their data values. Consider alternatives like cartograms, hex maps, or dot density maps when geographic area could mislead.

**5. Show your sources.**
Always include a data source caption. If the boundaries come from a specific year or administrative revision, say so.

**6. Consider interactivity.**
For web publication, `leaflet` or `mapview` let readers zoom and click. But a well-designed static map is often better for print and social sharing:

```{r}
#| label: interactive-map

library(leaflet)

# Quick interactive map
leaflet(breweries) |>
  addTiles() |>
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    radius = 2, color = "#C2001A",
    fillOpacity = 0.4, stroke = FALSE,
    popup = ~city
  )
```


## Hex bin maps: handling overplotting

When you have thousands of points, individual dots overlap. Hex bins aggregate into a grid while preserving the spatial pattern:

```{r}
#| eval: true
#| label: hexbin-map
#| fig-height: 6

ggplot() +
  geom_sf(data = germany, fill = NA, color = "gray40") +
  geom_hex(
    data = breweries,
    aes(x = lon, y = lat),
    bins = 30, alpha = 0.85
  ) +
  scale_fill_viridis_c(
    option = "inferno",
    direction = -1,
    name = "Breweries\nper hex"
  ) +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = c(0.15, 0.4)
  ) +
  labs(
    title = "Brewery density across Germany",
    subtitle = "Hexagonal bins aggregate point locations into a density surface",
    caption = "Source: biermap24.de"
  )
```

---

# Exercises {#exercises}

::: {.callout-note}
## Exercise 1: Brewery density per state

Calculate the **brewery density** (breweries per 1,000 km²) for each German state and create a choropleth map. Hint: use `st_area()` to get state areas. Which state has the highest density? Is it still Bavaria?
:::

::: {.callout-note}
## Exercise 2: Geocode and map

Pick 5–10 addresses of interest to you (e.g., Hertie School, Bundestag, your favorite restaurants) and geocode them using `tidygeocoder`. Plot them on a map of Berlin using `geom_sf()`. Bonus: add 1 km buffers around each point.
:::

::: {.callout-note}
## Exercise 3: Spatial filter

Using the brewery data, find all breweries within 100 km of Bamberg (the unofficial capital of Franconian beer). How many are there? Create a map showing the buffer and the breweries inside it.
:::

::: {.callout-note}
## Exercise 4: Critique a map

Find a map from a recent news article. Does it use an appropriate map type? Is the color scale honest? Could the geographic area bias be misleading? Write a short (3–4 sentence) critique.
:::

---

# Further reading

- Pebesma, E. & Bivand, R. (2023). [*Spatial Data Science: With Applications in R*](https://r-spatial.org/book/) — the definitive reference for `sf` and spatial analysis in R.
- Wilke, C.O. (2019). [*Fundamentals of Data Visualization*](https://clauswilke.com/dataviz/), Chapter 15: Visualizing geospatial data.
- Lovelace, R., Nowosad, J. & Muenchow, J. (2024). [*Geocomputation with R*](https://r.geocompx.org/) — more applied, great for geospatial workflows.
- The `tidygeocoder` [documentation](https://jessecambon.github.io/tidygeocoder/) for geocoding service details.

---

::: {.callout-note collapse="true"}
## Session info

```{r}
#| eval: true
sessionInfo()
```
:::
